\chapter{The Z-Wave Device API}
\label{c2}

This chapter describes the Z-Wave Device API and its use in detail All examples will use
the JSON REST API notation. Please not that the C library notation offers equal 
functionality but in a different notation.

The Z-Wave Device API is the north bound interface of the Z-Wave Core. This Z-Wave core
implement the whole control logic of the the Z-Wave network. The two main functions are
\begin{itemize}
\item Management of the network. This includes including and excluding devices, managing 
the routing and rerouting of the network  an executing some housekeeping functions 
to keep the network clean and stable. In the Z-Wave terminology all these functions 
are called 'function classes' and they are described in section \ref{c2:fc}. The function 
classes can be seen as functions offered by the controller itself. Hence the variables
and status parameters of the networks are offered by an object called 'controller'.

\item Execution of commands offered by the wireless devices as such switching switches 
and dimming dimmers. Z-Wave groups the command and their corresponding variables into
so called command classes. The Z-Wave API offers access to these command classes with
their variables and their commands according to the abilities od the respective device.
\end{itemize}
The next chapters first explain the timings of the communication in a wireless 
Z-Wave network. Then the data model is presented that reflects the real data and status 
information in the network.

The description of function Classes and command Classes and their access 
using the JSON API  complete the description of the Z-Wave Device API. For a full 
reference of function classes and command classes please refer to the Annex. 